The Castro AMR Simulation Code: Current and Future Developments

### 2020-01-31 astronum_2019_Castro/astronum-castro-2019.tex
update convergence

\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
\newcommand{\ddy}[1]{{\frac{{\partial#1}}{\partial y}}}
\newcommand{\ddz}[1]{{\frac{{\partial#1}}{\partial z}}}
\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
\newcommand{\Rbs}[1]{{\bf R} \left ( #1 \right )}
\newcommand{\Ic}{\bm{\mathcal{I}}}

@@

\castro\ implements both a second-order method (using a trapezoid rule
for the integral) and fourth-order method in space and time (using a
Simpsons rule for the integral and the spatial reconstruction of
\cite{mccorquodalecolella}).  We demonstrated that \castro\ achieves
the expected convergence on a wide variety of problems in
\cite{castro:sdc}.  At the moment, the method is limited to single
levels, but work is underway to extend this methodology to AMR.  For
problems where burning is important and can dominate the computational
expense, we hope this new SDC method will become the preferred
integration technique in the future.

>>>>

\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
\newcommand{\ddy}[1]{{\frac{{\partial#1}}{\partial y}}}
\newcommand{\ddz}[1]{{\frac{{\partial#1}}{\partial z}}}
\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}

\newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
\newcommand{\Rbs}[1]{{\bf R} \left ( #1 \right )}
\newcommand{\Ic}{\bm{\mathcal{I}}}

@@

\castro\ implements both a second-order method (using a trapezoid rule
for the integral) and fourth-order method in space and time (using a
Simpsons rule for the integral and the spatial reconstruction of
\cite{mccorquodalecolella}).  We demonstrated that \castro\ achieves
the expected convergence on a wide variety of problems in
\cite{castro:sdc}.  Table~\ref{table:convergence} summaries the
convergence of the standard unsplit CTU PPM hydrodynamics algorithm
with Strang splitting and the second- and fourth-order SDC methods on
a reacting acoustic pulse test problem.

At the moment, the method is limited to single
levels, but work is underway to extend this methodology to AMR.  For
problems where burning is important and can dominate the computational
expense, we hope this new SDC method will become the preferred
integration technique in the future.

\begin{table}[t]
\renewcommand{\arraystretch}{1.25}
\centering
\begin{tabular}{llllll}
{\bf field} & {$\epsilon_{64 \rightarrow 128}$} & 
{\bf rate} & {$\epsilon_{128\rightarrow 256}$} & 
{\bf rate} & {$\epsilon_{256\rightarrow 512}$} \\
\hline
\multicolumn{6}{c}{\bf Strang + CTU} \\
\hline
 $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.458 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
 $\rho X(\isotm{O}{16})$     & $1.717 \times 10^{14}$  & 1.648  & $5.479 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
\hline
\multicolumn{6}{c}{\bf 2nd order SDC} \\
\hline
 $\rho X(\isotm{C}{12})$     & $1.042 \times 10^{17}$  & 2.032  & $2.546 \times 10^{16}$  & 2.019  & $6.281 \times 10^{15}$  \\
 $\rho X(\isotm{O}{16})$     & $1.564 \times 10^{14}$  & 1.935  & $4.090 \times 10^{13}$  & 2.003  & $1.020 \times 10^{13}$  \\
\hline
\multicolumn{6}{c}{\bf 4th order SDC} \\
\hline
 $\rho X(\isotm{C}{12})$     & $8.789 \times 10^{15}$  & 3.798  & $6.319 \times 10^{14}$  & 3.911  & $4.201 \times 10^{13}$  \\
 $\rho X(\isotm{O}{16})$     & $1.294 \times 10^{13}$  & 3.765  & $9.518 \times 10^{11}$  & 3.872  & $6.501 \times 10^{10}$  \\
\hline
\end{tabular}
\renewcommand{\arraystretch}{1.0}
\caption{\label{table:convergence} Convergence of carbon mass density
  ($\rho X(\isotm{C}{12})$) and oxygen mass density ($\rho
  X(\isotm{O}{16})$) for the Strang-CTU algorithm and 2nd and 4th
  order SDC.  Data taken from \cite{castro:sdc}.}
\end{table}
